---
title: "Working with raster data - Tuesday pm"
author: "C. A. Hamm"
date: "June 21, 2016"
output: html_document
---

## Objetives
Start by putting up objectives/tasks that students will be working though:

1. Import a raster — A lidar canopy height model (lidar/Teak_lidarCHM.tif)

1. For the CHM, set values == 0 to NA (not trees)
Classify the raster according to some distribution – low medium and tall trees. This could be done using a histogram potentially or we could just decide that <2m is generally grasses / understory, <6m small trees,and the rest are tall trees. A function could import the desired thresholds. Visualize histogram/density and plot vertical cutoff lines.

1. PLOT - layer the classified raster, add a legend for each “class” - legends are super tricky to simplifying this process with a function would be good. see: http://neon-workwithdata.github.io/neon-data-institute-2016/R/classify-by-threshold-R/  for my take on forcing a legend outside of the plot area using par settings. You may have other better forms of magic to make this work well. :)

1. Export the plot figure to a pdf – publishable

1. Export the classified raster as a geotiff with NaFlag = -9999 to an outputs folder.

```{r housekeeping}
set.seed(9862144)
setwd("~/Desktop/Projects/NEON-2016/NEONDI-2016/Tuesday_pm/")
library("raster")
library("rgdal")
library("neonAOP")

# sink(paste0(format(Sys.time(), "%Y-%m-%d_%H%M%S"), "_sessionInfo.txt"))
# sessionInfo()
# sink()
sessionInfo()
```

### Objective \#1

```{r Objective_1}
# set path to file
TEAK_file <- "../NEONdata/D17-California/TEAK/2013/lidar/TEAK_lidarCHM.tif"
# import the file using the raster() function
tchm <- raster(TEAK_file)
```

### Objective \#2

```{r}
# set the 0 values to NA
tchm[tchm == 0] <- NA
# classify data based on some criterion
hist(tchm, col = "dodgerblue", breaks = 20, ylim = c(0, 1e4), ylab = "", xlab = "", main = "")
abline(v = c(2, 6, 15), lwd = 3, lty = 2, col = "red")

# will establish break ranges and convert to matrix
brackets <- matrix(c(0, 6, 1,
			6, 10, 2,
			10, 20, 3,
			20, 70, 4), 
			ncol = 3, byrow = TRUE)
brackets

re_tchm <- reclassify(tchm, brackets)
```
### Objective \#3 & \#4

```{r Objective_3}
# plot the figure
par(xpd = FALSE, mar = c(5, 4, 4, 6))
plot(re_tchm, col = c("light green", "dark green", "blue", "violet"), legend = FALSE)
par(xpd = TRUE)
legend(par()$usr[2], 4103300, legend = c("0 - 6m", "6 - 10m", "10 - 20m", "20 - 70m"), fill = c("light green", "dark green", "blue", "violet"), bty = "n", cex = 1.3)

# send the figure to a .pdf
pdf(file = "../outputs/TEAK/Teak_CHM_classified.pdf", bg = "white")
par(xpd = FALSE, mar = c(5, 4, 4, 6))
plot(re_tchm, col = c("light green", "dark green", "blue", "violet"), legend = FALSE)
par(xpd = TRUE)
legend(par()$usr[2], 4103300, legend = c("0 - 6m", "6 - 10m", "10 - 20m", "20 - 70m"), fill = c("light green", "dark green", "blue", "violet"), bty = "n", cex = 1.3)
dev.off()

```

### Objective \#5
```{r}
writeRaster(re_tchm, filename = "../outputs/TEAK/Teak_CHM_classified.tif", format = "GTiff", options = "COMPRESS = LZW", overwrite = TRUE, NAflag = -9999)
```